### 2004-05-13 Documentation/part2.tex
*** empty log message ***

% WW: CONTINUE HERE:

@@

Channels\index{Channel} represent an intermediate level in between 
the pure oscillation 
physics given by the oscillation probability $P_{\alpha\beta}$ and the
detected signal and background. A channel basically describes the path from one
initial state in the source to one interaction type (IT) in  the detector.  
Therefore a channel consists of the description of the initial state given 
by the polarity 
of the source, \eg\ muons or anti-muons in a neutrino factory, by the CP 
eigenvalue of the initial state, \ie\ neutrino or anti-neutrino and by the 
initial flavour. A channel furthermore consists of the interaction type in the
detector given by the final flavour of the neutrino, the interaction cross
section and the energy resolution function.

For the calculation of event rates, the first step is to compute the number of
events for each IT in the fiducial mass of the detector for each 
neutrino flavor and energy bin. The second step is to include the detector 
effects coming from
the insufficient knowledge used in the event reconstruction. We combine these
two steps in order to obtain the differential event rate spectrum for each
flavor and IT as seen by the detector, which we call a ``channel''.

The master formula for the differential event rate for each channel 
is given by
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{eqnarray}
\label{eq:master_event}
\frac{dn_{\beta}^{\text{IT}}}{dE'}=&&N\,\int \int dE\,d\hat{E}\quad
\underbrace{\Phi_{\alpha} (E)}_{\mathrm{Production}} \times \nonumber\\
&&\underbrace{\frac{1}{L^2} P_{(\alpha\rightarrow\beta)}(E,L,\rho;\theta_{23},
\theta_{12},\theta_{13},
\Delta m^2_{31},\Delta m^2_{21},\deltacp)}_{\mathrm{Propagation}}
\times \nonumber \\ &&\underbrace{\sigma^{\text{IT}}_f(E)
k_f^{\text{IT}}(E-\hat{E})}_{\mathrm{Interaction}} \times \nonumber \\
&&\underbrace{ T_f(\hat{E}) V_f(\hat{E}-E')}_{\mathrm{Detection}}\,,
\end{eqnarray}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
where $\alpha$ is the initial flavor of the neutrino and 
$\beta$ is the final flavour, 
$E$ is the incident neutrino energy, $\Phi_{\alpha} (E)$ is the flux of the 
initial flavor at the
source, $L$ is the baseline length, $N$ is a normalization factor, and 
$\rho$ is the matter density. The interaction term is composed of 
two factors, which are the total cross section 
$\sigma^{\text{IT}}_\beta(E)$ for the flavor $f$ and
the interaction type IT, and the energy distribution of the 
secondary particle $k_\beta^{\text{IT}}(E-\hat{E})$ with 
$\hat{E}$ the energy of the secondary particle. The detector properties are 
modeled by the threshold function $T_\beta(\hat{E})$, coming from the the 
limited resolution or the cuts in the analysis, and the energy resolution 
function $V_\beta(\hat{E}-E')$ of the secondary particle. Thus, $E'$ is the 
{\em reconstructed} neutrino energy.

Since it is rather cumbersome to numerically solve this double integral,
we split the two integrations. We first evaluate the integral over
$\hat{E}$, where the only terms containing $\hat{E}$ are
$k_\beta^{\text{IT}}(E-\hat{E})$,  $ T_\beta(\hat{E})$, and 
$ V_\beta(\hat{E}-E')$, and define:
\begin{eqnarray}
\label{eq:e_res} 
R_\beta^{\text{IT}}(E,E')\,\epsilon_\beta^{\text{IT}}(E')
 \equiv
\int d\hat{E} \quad T_\beta(\hat{E})\,k_\beta^{\text{IT}}(E-\hat{E})
\,V_\beta(\hat{E}-E')\,, 
e\end{eqnarray}

Thus $R_\beta^{\text{IT}}(E,E')$ describes the energy response of 
the detector, \ie\ a neutrino with a (true) energy of $E$ has a probability
of $R_\beta^{\text{IT}}(E,E') dE'$ to have a reconstructed energy 
in between $E'$ and $E'+dE'$. The function $R(E,E')$ is also refered to
as energy resolution function or smearing matrix in its actual representation
in the program\index{Energy resolution}\index{Smear matrix}. The detailed
definition and initialization of the energy resolution function is described
in section~\ref{sec:energy}.

The number of events in one channel and  in one bin\index{Bin} $i$ is given by
\begin{equation}
\label{eq:channel}
n_i^c=\int_{E_i-\Delta E/2}^{E_i+\Delta E/2} dE' \quad
\frac{dn_{\beta}^{\text{IT}}}{dE'} (E')
\end{equation}
where $c$ is the channel index.
Note that the events are binned according to their \emph{reconstructed} energy.
In order to perform the computation \GLOBES\ needs to know the flux,
the CP-sign of the initial state, the initial flavour, the final flavour,
the cross section and the energy resolution function. Employing the notation
of section~\ref{sec:energy} the full formula is

Thus the actual definition of a channel looks like
\begin{quote}
{\tt channel(\#channel\_1)<\\
\tb @channel = \#flux : $+$: muon: muon: \#cross: \#energy\\
>}
\end{quote}
Furthermore is is possible to define energy dependent efficiencies and
backgrounds (non-beam). These additional components can be either added
before the convolution in with $R(E,E')$ is done, \ie\ the integration with
respect to $E$ in equation~\ref{eq:e_res}. This is called 
{\tt @pre\_smearing\_$\ldots$}. The pre-smearing efficiencies and backgrounds
have to have as many elements as there are sampling points 
(see section~\ref{sec:energy}). In same cases the number of sampling points
is determined automatically by \GLOBES and the number of elements in the 
pre-smearing vectors has to be adjusted in the second run.
Of course, it is possible after all
integrations been performed to have similar factors  and this is called  
{\tt @post\_smearing\_$\ldots$}. All post-smearing factors have to have as 
many elements as there are bins. Efficiencies are multiplicative and their
default value is $1$ whereas backgrounds are additive and thus their default
value is $0$. Thus a more elaborate channel looks like
\begin{quote}
{\tt channel(\#channel\_1)<\\
\tb @channel = \#flux : $+$: muon: muon: \#cross: \#energy\\
\tb @pre\_smearing\_background = \{1,2,3,4,5,6,7,8,9,10\}\\
\tb @post\_smearing\_efficiencies = \{0.1,0.2,0.3,0.4,0.5\}\\
>}
\end{quote}
This experiment obviously uses $10$ sampling points and $5$ bins.

@@

@@

In order to obtain the number of events 
$n_i^c$ from one channel $c$ in one bin $i$ one has to solve this integral
%
\begin{eqnarray}
\label{eq:events_bin}
n_i^c=N/L^2\,\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
\quad \int dE \,\, \Phi^c(E)\,
P^c(E)\,
\sigma^c(E)\,
R^c(E,E')\,
\epsilon^c(E')\,.
\end{eqnarray} 
%

@@

@@

@@

>>>>

@@

Channels\index{Channel} in \GLOBES\ represent an intermediate level 
between the pure oscillation physics given by the oscillation probability
$P_{\alpha\beta}$, and the total event rates composed of signal and 
background. A channel describes the path from one
initial neutrino flavor in the source to the event rates in the detector for one specific interaction type (IT) and final flavor.  
Therefore, a channel contains the description of the 
initial neutrino flavor, its CP eigenvalue (neutrino or antineutrino), 
the detected neutrino flavor, the interaction cross sections for the chosen interaction type, and the energy resolution function of the detector.

Before we come to the definition of the channel in AEDL, we introduce the general concept for the calculation of event rates. The first step is to
compute the number of events for each IT in the detector for each 
initial and final neutrino flavor and energy bin. The second step is to include the detector effects coming from the insufficient knowledge in the event reconstruction. These two steps combined lead to the differential event rate spectrum for each initial and final flavor and IT as seen by the detector, which we call the ``channel''. In this section, we focus on the
 first step, \ie , we discuss the definition of the energy resolution function in the next section, since this is a rather comprehensive issue.

The differential event rate for each channel is given by
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{eqnarray}
\label{eq:master_event}
\frac{dn_{\beta}^{\text{IT}}}{dE'}=&&N\,\int\limits_0^\infty \int\limits_0^\infty dE\,d\hat{E}\quad
\underbrace{\Phi_{\alpha} (E)}_{\mathrm{Production}} \times \nonumber\\
&&\underbrace{\frac{1}{L^2} P_{(\alpha\rightarrow\beta)}(E,L,\rho;\theta_{23},
\theta_{12},\theta_{13},
\Delta m^2_{31},\Delta m^2_{21},\deltacp)}_{\mathrm{Propagation}}
\times \nonumber \\ &&\underbrace{\sigma^{\text{IT}}_f(E)
k_f^{\text{IT}}(E-\hat{E})}_{\mathrm{Interaction}} \times \nonumber \\
&&\underbrace{ T_f(\hat{E}) V_f(\hat{E}-E')}_{\mathrm{Detection}}\,,
\end{eqnarray}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
where $\alpha$ is the initial flavor of the neutrino, 
$\beta$ is the final flavor, $\Phi_{\alpha} (E)$ is the flux of the 
initial flavor at the
source, $L$ is the baseline length, $N$ is a normalization factor, and 
$\rho$ is the matter density. The energies in this formula are given as follows:
\begin{itemize}
\item
 $E$ is the incident neutrino energy, \ie, the actual energy of the 
incoming neutrino (which is not directly accessible to the experiment)
\item
 $\hat{E}$ is the energy of the secondary particle
\item
 $E'$ is the reconstructed neutrino energy, \ie, the measured
neutrino energy as obtained from the experiment
\end{itemize}
The interaction term is composed of 
two factors, which are the total cross section 
$\sigma^{\text{IT}}_\beta(E)$ for the flavor $f$ and
the interaction type IT, and the energy distribution of the 
secondary particle $k_\beta^{\text{IT}}(E-\hat{E})$.
The detector properties are 
modeled by the threshold function $T_\beta(\hat{E})$, coming from the the 
limited resolution or the cuts in the analysis, and the energy resolution 
function $V_\beta(\hat{E}-E')$ of the secondary particle.

Since it is a lot of effort to solve this double integral numerically,
we split up the two integrations. First, we evaluate the integral over
$\hat{E}$, where the only terms containing $\hat{E}$ are
$k_\beta^{\text{IT}}(E-\hat{E})$,  $ T_\beta(\hat{E})$, and 
$ V_\beta(\hat{E}-E')$. We define:
\begin{eqnarray}
\label{eq:e_res} 
R_\beta^{\text{IT}}(E,E')\,\epsilon_\beta^{\text{IT}}(E')
 \equiv
\int\limits_0^\infty d\hat{E} \quad T_\beta(\hat{E})\,k_\beta^{\text{IT}}(E-\hat{E})
\,V_\beta(\hat{E}-E')\,. 
\end{eqnarray}
Thus, $R_\beta^{\text{IT}}(E,E')$ describes the energy response of 
the detector, \ie , a neutrino with a (true) energy $E$ is reconstructed
with an energy between $E'$ and $E'+dE'$ with a probability
$R_\beta^{\text{IT}}(E,E') dE'$. The function $R(E,E')$ is also often called ``energy resolution function''. Actually, its internal representation
in the software is a smearing matrix\index{Energy resolution}\index{Smear matrix}. The function $\epsilon_\beta^{\text{IT}}(E')$ will later be refered to as ``post-smearing efficiencies'', since it will allow us to define cuts and threshold functions {\em after} the smearing is performed, \ie, as function of $E'$. The detailed definition and initialization of the energy resolution function is described in \Sec~\ref{sec:energy}.

Eventually, we can write down the number of events per bin $i$ \index{Bin} and channel $c$ as
\begin{equation}
\label{eq:channel}
n_i^c=\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' \quad
\frac{dn_{\beta}^{\text{IT}}}{dE'} (E').
\end{equation}
This means that one has to solve this integral
\begin{eqnarray}
\label{eq:events_bin}
n_i^c=N/L^2\,\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
\quad \int\limits_0^\infty dE \,\, \Phi^c(E)\,
P^c(E)\,
\sigma^c(E)\,
R^c(E,E')\,
\epsilon^c(E')\,.
\end{eqnarray} 
Note that the events are binned according to their \emph{reconstructed} energy.

A simple channel definition in \GLOBES\ consists of the flux,
the CP-sign of the initial state, the initial flavor, the final flavor,
the cross sections, and the energy resolution function. In order to refer to
flux, cross sections, and energy resolution functions, they have to be 
defined before with their {\tt \#name} in the respective environments. 
Thus, a simple definition of a channel is
\begin{quote}
{\tt channel(\#channel\_1)<\\
\tb @channel = \#flux : $+$: muon: muon: \#cross: \#energy\\
>}
\end{quote}
Note that the energy environment will be described in the next section.

% WW: HIER WEITER

@@

The definition of the energy resolution function is rather sophisticated in 
\GLOBES . In particular, since this aspect is rather time-critical, several
algorithms are implemented. The choice of the proper algorithm depends on the
experiment and the frequencies of the involved neutrino oscillations.

In this section, we first discuss the principles of the energy smearing, where it is assumed that the reader is familiar with \Sec~\ref{sec:channel}. Then we introduce an automatic energy smearing algorithm, which is fairly simple to understand and applicable to most beam-based experiments. 
In most cases, the reader may want to continue to the next section after reading these two subsections. In the third subsection, we describe a more elaborate (and slower) smearing algorithm, which can be used together with rather fast neutrino oscillations compared to the bin size, such as for reactor experiments. Eventually, we show how one can use a manual smearing matrix instead of using one of the implemented algorithms.

\subsection{Introduction and principles}

\begin{figure}[t!]
\begin{center}
\includegraphics[width=15cm]{energies}
\end{center}
\caption{\label{fig:energies} The different evaluation levels for the energy smearing in \GLOBES .}
\end{figure}

DESCRIBE PICTURE!

@@

@@

ADJUST TO SECTION AND REPLACE:

In combination with the channels, is is possible to define energy 
dependent efficiencies and (non-beam) backgrounds. Efficiencies are multiplicative factors, whereas backgrounds are added to the event rates. These components can be introduced before or after the convolution in \eq~(\ref{eq:e_res}) is done. If they are introduced before, we call them
{\tt @pre\_smearing\_efficiencies} or {\tt @pre\_smearing\_background}. 
If they are introduced after, we call them {\tt @post\_smearing\_efficiencies} or {\tt @post\_smearing\_background}.
Note that pre-smearing components are always afunction of the incident neutrino energy $E$. Thus, there have to be as  many elements as there are sampling points (see \Sec~\ref{sec:energy}). In same cases, the number of sampling points is determined automatically by \GLOBES , and the number of elements in the  pre-smearing vectors has to be adjusted in the second run.
Examples for pre-smearing quantities are non-beam backgrounds, such as from geophysical neutrinos. The post-smearing components are always a function of the reconstructed neutrino energy $E'$, such as the post-smearing efficiencies $\epsilon_\beta^{\text{IT}}(E')$ in \eq~(\ref{eq:e_res}). Examples for post-smearing efficiencies are cuts and detection threshold functions. All post-smearing components have to have as 
many elements as there are energy bins. Efficiencies are multiplicative 
and their default value is $1$, whereas backgrounds are additive and their default value is $0$. Thus, a more elaborate channel can be defined as
\begin{quote}
{\tt channel(\#channel\_1)<\\
\tb @channel = \#flux : $+$: muon: muon: \#cross: \#energy\\
\tb @pre\_smearing\_background = \{1,2,3,4,5,6,7,8,9,10\}\\
\tb @post\_smearing\_efficiencies = \{0.1,0.2,0.3,0.4,0.5\}\\
>}
\end{quote}
This experiment uses $10$ sampling points and $5$ bins.

\subsection{Bin based automatic energy smearing}

@@

\subsection{Low-pass filter based automatic energy smearing}

@@

\subsection{Manual energy smearing}
